Mining of Keystroke and Mouse Dynamics to Enhance Student Engagement with Programming Assignments

### 2018-04-18 paper.tex
New Title
New abstract for #7
@JJ please check

% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Enhancing Student Engagement via Reduction of Frustration with Programming Assignments using Machine Learning}

@@

\keywords{Affective computing, neural networks, learning analytics}

\abstract{Learning to program is often regarded as a difficult task. When
selecting an appropriate programming assignment, experienced instructors
gauge a student\'s affective state and skills to then assign an
activity with the appropriate level of difficulty. This work is
focused on the prediction of the affective states of programmers with
different levels of expertise when learning a new programming
language. For this, an interactive web-based programming platform is
proposed. The platform is designed to collect data from the students\'
interaction for data analysis. Current work is focused on the
prediction of affective states using non-obtrusive
sensors. Specifically, the aim of this research is to evaluate the use
of keyboard and mouse dynamics as an appropriate sensory input for an
affective recognition system. The proposed method uses feature vectors
obtained by mining data generated from both keyboard and mouse
dynamics of students as they work in basic Python programming
assignments, which were used to train different classification
algorithms to classify learners into five different affective states:
boredom, frustration, distraction, relaxation and engagement. Accuracy
achieved was around 75\% with J48 obtaining the best results, proving
that data gathered from non-obtrusive sensors can successfully be used
as another input to classification models in order to predict an
individual\'s affective states.}

>>>>

% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Mining of Keystroke and Mouse Dynamics to Enhance Student Engagement with Programming Assignments}

@@

\keywords{Affective computing, neural networks, machine learning}

\abstract{
When choosing from a collection of programming assignments, instructors must
gauge the student\'s affective state and skills to select a learning resource
with the appropriate level of difficulty. This work is
focused on capturing and pre-processing of data for the evaluation of several
machine learning techniques aim at classifing the affective
states of users with different levels of expertise
when learning a new programming language.
Currently, web-based platform is applied to collect data from the students\'
interactions whith programming excersises. The aim of this research is to
evaluate the use of keyboard and mouse dynamics as an appropriate
non-obtrusive sensory input for an affective aware system.
Feature vectors were obtained by pre-processing
data from keyboard and mouse dynamics captured from students as
they were coding basic Python programming assignments.
Feature vectors were used for the classification of five affective states:
boredom, frustration, distraction, relaxation and engagement.
The following classification algorithms have been evaluated: k -nearest
neighbors,feed forward neural networks, na\"ive Bayes classifier and the J-48
tree indiction algorithm). The best accuracy was around 75\% and was achieved
by the J-48 algorithm. Results show that data gathered from ready-available,
non-obtrusive sensors can be used successfully
as another input to hyebrid classification models in order to predict an
individual\'s affective states.}
